---
title: "Time Series Processing"
author: "Romana Limberger, Uriah Daugaard et al."
output:
  html_document:
    number_sections: true
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Dependencies, functions and data

## Dependencies

```{r packages, message=FALSE}
library(tidyverse)
library(randomForest)
library(data.table)
library(pracma)
library(parallel)
library(doParallel)
library(here)
library(segmented)

set.seed(1)
options(dplyr.summarise.inform = FALSE)
```

## Functions

```{r load_own_fcts}
source(file = here("R/1_Functions/OtherFunctions.R")) ## other useful functions
```


## Data {#data}

```{r data, warning=FALSE, message=FALSE}
# Species densities
densities <- read_csv("../../Data/2_TimeSeriesData/Untransformed/AbundanceTimeSeries.csv")
# Oxygen 
o2 <- read_csv("../../Data/2_TimeSeriesData/Untransformed/OxygenTimeSeries.csv")
# Water chemistry
water_chem <- read_csv("../../Data/2_TimeSeriesData/Untransformed/CarbonNitrogenTimeSeries.csv")

days <- unique(densities$day)
timestamps <- unique(densities$timestamp)
days_timestamps <- data.frame(day=days, timestamp=timestamps)
```


# Preliminary data processing

## Taking the mean of repeated measurements

```{r taking_the_mean, message=FALSE, warning=FALSE}
densities$method <- ifelse(grepl("bemovi", densities$measurement, fixed = TRUE),"Video",
                           ifelse(densities$measurement=="manualcount", "Manual count", 
                                  ifelse(densities$measurement=="flowcam","FlowCam","Flowcytometer")))

densities <- densities %>%
  group_by(timestamp, day, bottle, species, light_treatment, richness, composition, incubator, method) %>%
  summarize(density=mean(density, na.rm = T),
            biomass=mean(biomass, na.rm = T)) %>%
  mutate(density = ifelse(is.nan(density),as.numeric(NA),density),
         biomass = ifelse(is.nan(biomass),as.numeric(NA),biomass))

o2_bothMeas <- o2

o2 <- o2 %>%
  group_by(timestamp, day, bottle, richness, composition, incubator) %>%
  summarize(percent_o2=ifelse(length(percent_o2)==1,NA,mean(percent_o2, na.rm = T))) %>%
  mutate(percent_o2 = ifelse(is.nan(percent_o2),as.numeric(NA),percent_o2))
```


## Marking each time sereis as either "forecast target" or "non-forecast target"

```{r, message=FALSE, warning=FALSE}
colpi <- data.frame(species = "Colpidium striatum",
                    bottle = c(17,20,27))
eupl <- data.frame(species = "Euplotes daidaleos",
                   bottle = c(18))
dex <- data.frame(species = "Dexiostoma campylum",
                  bottle = c(23))
lox <- data.frame(species = "Loxocephalus sp.",
                  bottle = c(25))
cosm <- data.frame(species = "Cosmarium botrytis",
                   bottle = c(1,3,5,6,11,18,19,22,23,24,27,29))
sty <- data.frame(species = "Stylonychia sp.",
                  bottle=c(27))
stau2 <- data.frame(species = "Staurastrum polytrichum",
                    bottle = c(8,10,21,22,27,30))
stau1 <- data.frame(species = "Staurastrum gracile",
                    bottle = c(6,12,19,20,28,29))
desm <- data.frame(species = "Desmodesmus armatus",
                   bottle = c(2))
desm_clump <- data.frame(species = "Desmodesmus clumps",
                         bottle = c(2))

speciesKeptButNotForecasts <- rbind(colpi,eupl,dex,lox,cosm,sty,stau2,stau1,desm,desm_clump)

speciesKeptButNotForecasts$bottle <- ifelse(speciesKeptButNotForecasts$bottle<10,
                                            paste0("b_0",speciesKeptButNotForecasts$bottle),
                                            paste0("b_",speciesKeptButNotForecasts$bottle))

densities <- full_join(densities, speciesKeptButNotForecasts)

speciesKeptButNotForecasts$int <- interaction(speciesKeptButNotForecasts$bottle, speciesKeptButNotForecasts$species)

targets <- c("Bacteria","Chlamydomonas clumps","Chlamydomonas reinhardtii", "Coleps sp.","Cosmarium botrytis",
             "Desmodesmus armatus","Desmodesmus clumps","Dividing Chlamydomonas","Paramecium caudatum","Paramecium bursaria","Small cells",
             "Staurastrum gracile","Monoraphidium obtusum","Staurastrum polytrichum","Stylonychia sp.","Colpidium striatum",
             "Dexiostoma campylum","Euplotes daidaleos","Loxocephalus sp.")
other <- c("Debris", "Didinium nasutum", "Digested algae", "Small unidentified","Colpidium vacuoles")

densities <- densities %>%
  dplyr::mutate(bottle.species = interaction(bottle, species)) %>%
  dplyr::mutate(ForecastTarget = ifelse(!(bottle.species %in% speciesKeptButNotForecasts$int) & species %in% targets,
                                        T,F)) %>%
  dplyr::select(-bottle.species)
```


## Calculating realized richness and community biomass
```{r}
vars <- c("Coleps sp.","Paramecium bursaria","Paramecium caudatum", 
          "Stylonychia sp.","Colpidium striatum","Euplotes daidaleos",
          "Dexiostoma campylum","Loxocephalus sp.","Chlamydomonas reinhardtii",
          "Cosmarium botrytis",
          "Desmodesmus armatus", "Staurastrum gracile","Monoraphidium obtusum",
          "Staurastrum polytrichum","Bacteria")  

realized.richness <- densities %>%
  dplyr::filter(species %in% vars, ForecastTarget==T) %>%
  group_by(bottle,richness,composition,light_treatment) %>%
  dplyr::summarize(realized_richness = length(unique(species)))

biomass <- densities %>%
  group_by(timestamp, day, bottle, light_treatment, richness, composition, incubator) %>%
  summarise(biomass = sum(biomass, na.rm = T))
```

## Addition of missing data as NA (and addition of realized richness)

```{r, message=FALSE, warning=FALSE}
# biomass
biomass$timestamp <- as.numeric(biomass$timestamp)

biomass.temp <- biomass %>%
  dplyr::filter(timestamp == 20220629) %>%
  mutate(timestamp = 20220701,
         day = 284, 
         biomass = NA)

biomass <- full_join(biomass, biomass.temp)
rm(biomass.temp)

# densities & biomass
densities$timestamp <- as.numeric(densities$timestamp)

dens.temp <- densities %>%
  dplyr::filter(timestamp == 20220629) %>%
  mutate(timestamp = 20220701,
         day = 284, 
         density = NA,
         biomass = NA)

densities <- full_join(densities, dens.temp)
rm(dens.temp)

## bacteria
bac <- densities %>% dplyr::filter(species=="Bacteria")
grid <- expand_grid(bottle=unique(bac$bottle), day=days)
grid <- full_join(grid, days_timestamps)

bac_structure <- bac %>%
  group_by(bottle,species,light_treatment,richness,composition,incubator,method,ForecastTarget) %>%
  summarise(n=n())%>%
  dplyr::select(-n)

grid <- full_join(grid, bac_structure)
bac <- full_join(bac,grid %>% mutate(timestamp=as.numeric(timestamp)))
densities <- full_join(densities, bac)

## algae

Algae <- densities %>% dplyr::filter(species=="Algae")
grid <- expand_grid(bottle=unique(Algae$bottle), day=days)
grid <- full_join(grid, days_timestamps)

Algae_structure <- Algae %>%
  group_by(bottle,species,light_treatment,richness,composition,incubator,method,ForecastTarget) %>%
  summarise(n=n())%>%
  dplyr::select(-n)

grid <- full_join(grid, Algae_structure)
Algae <- full_join(Algae,grid %>% mutate(timestamp=as.numeric(timestamp)))
densities <- full_join(densities, Algae)

## Didinium

grid$species <- "Didinium nasutum"
grid$method <- "Manual count"
grid$ForecastTarget <- F
did <- densities %>% dplyr::filter(species=="Didinium nasutum")

did <- full_join(did,grid%>% mutate(timestamp=as.numeric(timestamp),
                                    day=as.double(day)), by=colnames(grid))

densities <- full_join(densities, did)

densities <- densities %>% #last day of Didinium set to 0 for interpolation
  mutate(density = ifelse(species=="Didinium nasutum" & day==287, 0, density),
         biomass = ifelse(species=="Didinium nasutum" & day==287, 0, biomass))

# Carbon and Nitrogen

water_chem.temp <- water_chem %>%
  dplyr::filter(timestamp == 20220704) %>%
  mutate(timestamp = 20220701,
         day = 284, 
         concentration = NA)

water_chem <- full_join(water_chem, water_chem.temp)
rm(water_chem.temp)

grid <- expand_grid(bottle=unique(water_chem$bottle),
                    day=days,
                    type=unique(water_chem$type))

grid <- full_join(grid, days_timestamps)
chem_structure <- water_chem %>%
  group_by(bottle,type,light_treatment,richness,composition,incubator) %>%
  summarise(n=n())%>%
  dplyr::select(-n)

grid <- full_join(grid, chem_structure)
water_chem <- full_join(water_chem,grid %>% mutate(timestamp=as.numeric(timestamp)))

# Oxygen

o2.temp <- o2 %>%
  dplyr::filter(timestamp == 20220704) %>%
  mutate(timestamp = 20220701,
         day = 284, 
         percent_o2 = NA)

o2 <- full_join(o2, o2.temp)

o2.temp <- o2_bothMeas %>%
  dplyr::filter(timestamp == 20220704) %>%
  mutate(timestamp = 20220701,
         day = 284, 
         percent_o2 = NA)

o2_bothMeas <- full_join(o2_bothMeas, o2.temp)

o2.temp <- o2 %>%
  dplyr::filter(timestamp == 20220704) %>%
  mutate(timestamp = 20211201,
         day = 72, 
         percent_o2 = NA)

o2 <- full_join(o2, o2.temp)

o2.temp <- o2_bothMeas %>%
  dplyr::filter(timestamp == 20220704) %>%
  mutate(timestamp = 20211201,
         day = 72, 
         percent_o2 = NA)

o2_bothMeas <- full_join(o2_bothMeas, o2.temp)
rm(o2.temp)

## Adding realized richness
biomass <- full_join(biomass, realized.richness)
densities <- full_join(densities, realized.richness)
o2 <- full_join(o2, realized.richness)
o2_bothMeas <- full_join(o2_bothMeas, realized.richness)
water_chem <- full_join(water_chem, realized.richness)
```


# Interpolation, detrending and transformation of data {#interpolation}

## Interpolation and imputation (creation of equidistant data):

```{r, message=FALSE, warning=FALSE}
## impute and interpolate

# Total biomass

time.interp <- time.interp.fun(timestep = biomass$day)
biomass.int <- biomass %>% na.omit() %>% arrange(day) 
biomass.int <- 
  setDT(biomass.int)[, list(day = time.interp,
                            biomass = chi(day, biomass, time.interp)), 
                     by = .(bottle,composition,
                            light_treatment,
                            richness,incubator, realized_richness)]

# species densities and biomass
time.interp <- time.interp.fun(timestep = densities$day)
densities.int <- densities %>% na.omit() %>% arrange(day)  %>%
  dplyr::filter(species !="Didinium nasutum")
densities.int <- 
  setDT(densities.int)[, list(day = time.interp,
                              density = chi(day, density, time.interp),
                              biomass = chi(day, biomass, time.interp)), 
                       by = .(species,bottle,composition,
                              light_treatment,ForecastTarget,
                              richness,incubator,method, realized_richness)]

densities.int$density <- ifelse(densities.int$density<0,0,densities.int$density)
densities.int$biomass <- ifelse(densities.int$biomass<0,0,densities.int$biomass)

densities.int.other <- densities %>%
  dplyr::filter(species %in% other) %>%
  drop_na("density") %>%
  arrange(day)

densities.int.other <- 
  setDT(densities.int.other)[, list(day = time.interp,
                                    density = chi(day, density, time.interp)),
                             by = .(species,bottle,composition,
                                    light_treatment,ForecastTarget,
                                    richness,incubator,method, realized_richness)]

densities.int.other$density <- ifelse(densities.int.other$density<0,0,densities.int.other$density)
densities.int.other$biomass <- NA

densities.int <- full_join(densities.int, densities.int.other)

# Oxygen

time.interp <- time.interp.fun(timestep = o2$day)
o2.int <- o2 %>% na.omit() %>% arrange(day)
o2.int$method <- "o2meter"
o2.int <- setDT(o2.int)[, list(day = time.interp,
                               percent_o2 = chi(day, percent_o2, time.interp)), 
                        by = .(bottle,composition,
                               light_treatment, richness,
                               incubator,method, realized_richness)]


# Water chemistry

time.interp <- time.interp.fun(timestep = water_chem$day)
water_chem.int <- water_chem %>% dplyr::select(-cv, -n) %>% na.omit() %>% arrange(day)
water_chem.int <- setDT(water_chem.int)[, list(day = time.interp,
                                               concentration = chi(day, concentration, time.interp)),
                                        by = .(bottle,composition,
                                               light_treatment, richness,
                                               incubator,type, realized_richness)]
```


## Detrending & standardization

```{r, message=FALSE, warning=FALSE}
## remove linear trend

## total biomass

biomass.int.detrend.complete <- setDT(biomass.int)[, list(day = day,
                                                          biomass.original = biomass,
                                                          biomass.seg = resid(segmented(lm(biomass~day), seg.Z = ~day, psi=c(93,198))),
                                                          biomass = residuals(lm(biomass~day))),
                                                   by = .(bottle,richness,
                                                          composition,incubator,
                                                          light_treatment, realized_richness)]

## species densities and biomass

DesmoB2.int <- densities.int %>%
  filter(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02")

densities.int.detrend.complete <- densities.int %>%
  filter(!(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02"))

densities.int.detrend.complete <- 
  setDT(densities.int.detrend.complete)[, list(day = day,
                                               density.original = density, # original density TS
                                               biomass.original = biomass, # original biomass TS
                                               density.seg = resids_seg_lm(density, day, c(93,198)), # density detrended segmented
                                               biomass.seg = resids_seg_lm(biomass, day, c(93,198)), # biomass detrended segmented
                                               density = residuals(lm(density~day)), # density detrended linear simple
                                               biomass = resids_lm(biomass, day)), # biomass detrended linaer simple
                                        by = .(species,bottle,richness,
                                               composition,incubator,
                                               method,light_treatment,ForecastTarget,
                                               realized_richness)]

DesmoB2.int.detrend.complete <- setDT(DesmoB2.int)[, list(day = day,
                                                          density.original = density,
                                                          biomass.original = biomass,
                                                          density.seg = resids_seg_lm(density, day, c(93,198)),
                                                          biomass.seg = resids_seg_lm(biomass, day, c(93,198)),
                                                          density = residuals(lm(density~day)),
                                                          biomass = resids_lm(biomass, day)),
                                                   by = .(species,bottle,richness,
                                                          composition,incubator,
                                                          method,light_treatment,ForecastTarget,
                                                          realized_richness)]

## Oxygen

o2.int.detrend.complete <- setDT(o2.int)[, list(day = day,
                                                percent_o2.original = percent_o2,
                                                percent_o2.seg = resid(segmented(lm(percent_o2~day), seg.Z = ~day, psi=c(93,198))),
                                                percent_o2 = residuals(lm(percent_o2~day))),
                                         by = .(bottle,
                                                composition, richness,incubator,
                                                method,light_treatment, realized_richness)]

## water chemistry

water_chem.int.detrend.complete <- setDT(water_chem.int)[, list(day = day,
                                                                concentration.original = concentration,
                                                                concentration.seg = resid(segmented(lm(concentration~day), 
                                                                                                    seg.Z = ~day, psi=c(93,198))),
                                                                concentration = residuals(lm(concentration~day))),
                                                         by = .(type,bottle,
                                                                composition, richness,incubator,
                                                                light_treatment, realized_richness)]

##############################################################
## Standardization

biomass.int.detrend.complete <- biomass.int.detrend.complete %>%
  group_by(bottle, richness, light_treatment, composition, incubator, realized_richness) %>%
  mutate(biomass.detrended.unstandardized = biomass + mean(biomass.original),
         biomass.seg.detrended.unstandardized = biomass.seg + mean(biomass.original),
         biomass = biomass/sd(biomass, na.rm = T),
         biomass.seg = biomass.seg/sd(biomass.seg, na.rm = T))


densities.int.detrend.complete <- densities.int.detrend.complete %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  mutate(density.detrended.unstandardized = density + mean(density.original),
         biomass.detrended.unstandardized = biomass + mean(biomass.original),
         density.seg.detrended.unstandardized = density.seg + mean(density.original),
         biomass.seg.detrended.unstandardized = biomass.seg + mean(biomass.original), 
         density = density/sd(density, na.rm = T), 
         biomass = biomass/sd(biomass, na.rm = T), 
         density.seg = density.seg/sd(density.seg, na.rm = T), 
         biomass.seg = biomass.seg/sd(biomass.seg, na.rm = T)) %>% 
  mutate(density = ifelse(is.nan(density),0,density), # because didinium
         biomass = ifelse(is.nan(biomass),0,biomass),
         density.seg = ifelse(is.nan(density.seg),0,density.seg), # because didinium
         biomass.seg = ifelse(is.nan(biomass.seg),0,biomass.seg))


DesmoB2.int.detrend.complete <- DesmoB2.int.detrend.complete %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  mutate(density.detrended.unstandardized = density + mean(density.original),
         biomass.detrended.unstandardized = biomass + mean(biomass.original),
         density.seg.detrended.unstandardized = density.seg + mean(density.original),
         biomass.seg.detrended.unstandardized = biomass.seg + mean(biomass.original), 
         density = density/sd(density, na.rm = T), 
         biomass = biomass/sd(biomass, na.rm = T), 
         density.seg = density.seg/sd(density.seg, na.rm = T), 
         biomass.seg = biomass.seg/sd(biomass.seg, na.rm = T)) 


densities.int.detrend.complete <- rbind(densities.int.detrend.complete,
                                        DesmoB2.int.detrend.complete)


o2.int.detrend.complete <- o2.int.detrend.complete %>%
  group_by(bottle, light_treatment, composition, incubator, method, richness, realized_richness) %>%
  mutate(percent_o2.detrended.unstandardized = percent_o2 + mean(percent_o2.original),
         percent_o2.seg.detrended.unstandardized = percent_o2.seg + mean(percent_o2.original),
         percent_o2 = percent_o2/sd(percent_o2, na.rm = T),
         percent_o2.seg = percent_o2.seg/sd(percent_o2.seg, na.rm = T))

water_chem.int.detrend.complete <- water_chem.int.detrend.complete %>%
  group_by(bottle, light_treatment, composition, incubator, type, richness, realized_richness) %>%
  mutate(concentration.detrended.unstandardized = concentration + mean(concentration.original),
         concentration.seg.detrended.unstandardized = concentration.seg + mean(concentration.original),
         concentration = concentration/sd(concentration, na.rm = T),
         concentration.seg = concentration.seg/sd(concentration.seg, na.rm = T))
```

# Further processing/calculations

## Adding more information regarding the light treatment.

```{r}
# adding light schedule
dd_list <- list(biomass = biomass, 
                biomass.int.detrend.complete = biomass.int.detrend.complete, 
                densities = densities, 
                densities.int.detrend.complete = densities.int.detrend.complete, 
                o2 = o2, 
                o2.int.detrend.complete = o2.int.detrend.complete, 
                water_chem = water_chem,
                water_chem.int.detrend.complete = water_chem.int.detrend.complete)

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}

dd_list <- mclapply(dd_list, function(df){
  df <- df %>%
    dplyr::mutate(light = case_when(light_treatment =="constant" ~ 30,
                                    day >= 191 ~ 1,
                                    184 <= day & day < 191 ~ 2,
                                    177 <= day & day < 184 ~ 4,
                                    170 <= day & day < 177 ~ 6,
                                    163 <= day & day < 170 ~ 8,
                                    156 <= day & day < 163 ~ 10,
                                    149 <= day & day < 156 ~ 12,
                                    142 <= day & day < 149 ~ 14,
                                    135 <= day & day < 142 ~ 16,
                                    128 <= day & day < 135 ~ 18,
                                    121 <= day & day < 128 ~ 20,
                                    114 <= day & day < 121 ~ 22,
                                    107 <= day & day < 114 ~ 24,
                                    100 <= day & day < 107 ~ 26,
                                    93 <= day & day < 100 ~ 28,
                                    day < 93 ~ 30),
                  light_phase = case_when(light_treatment =="constant" ~ "constant",
                                          day < 93 ~ "constant before decrease",
                                          day >= 93 & day < 198 ~ "decreasing",
                                          day >= 198 ~ "constant after decrease"))
  df
}, mc.cores = cr)
```

```{r}
# adding light schedule
dd_list.int <- list(biomass.int = biomass.int, 
                    densities.int = densities.int, 
                    o2.int = o2.int, 
                    water_chem.int = water_chem.int)

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}

dd_list.int <- mclapply(dd_list.int, function(df){
  df <- df %>%
    dplyr::mutate(light = case_when(light_treatment =="constant" ~ 30,
                                    day >= 191 ~ 1,
                                    184 <= day & day < 191 ~ 2,
                                    177 <= day & day < 184 ~ 4,
                                    170 <= day & day < 177 ~ 6,
                                    163 <= day & day < 170 ~ 8,
                                    156 <= day & day < 163 ~ 10,
                                    149 <= day & day < 156 ~ 12,
                                    142 <= day & day < 149 ~ 14,
                                    135 <= day & day < 142 ~ 16,
                                    128 <= day & day < 135 ~ 18,
                                    121 <= day & day < 128 ~ 20,
                                    114 <= day & day < 121 ~ 22,
                                    107 <= day & day < 114 ~ 24,
                                    100 <= day & day < 107 ~ 26,
                                    93 <= day & day < 100 ~ 28,
                                    day < 93 ~ 30),
                  light_phase = case_when(light_treatment =="constant" ~ "constant",
                                          day < 93 ~ "constant before decrease",
                                          day >= 93 & day < 198 ~ "decreasing",
                                          day >= 198 ~ "constant after decrease"))
  df
}, mc.cores = cr)
```


## Calculating Daily species richness and their average (mean & median)

```{r, message=FALSE}
# raw
day_spec_rich <- dd_list$densities %>%
  dplyr::filter(species %in% vars, 
                (method=="FlowCam" & density>=100) | (method=="Video" & density>=10) | (method=="Flowcytometer")) %>%
  group_by(bottle, day) %>%
  summarise(day_richness = n()) %>%
  dplyr::mutate(day_richness = ifelse(day==284,NA,day_richness)) %>%
  group_by(bottle) %>%
  dplyr::mutate(mean_richness = mean(day_richness, na.rm = T),
                median_richness = median(day_richness, na.rm = T))

dd_list$densities <- full_join(dd_list$densities, day_spec_rich)
dd_list$biomass <- full_join(dd_list$biomass, day_spec_rich)
dd_list$o2 <- full_join(dd_list$o2, day_spec_rich)
dd_list$water_chem <- full_join(dd_list$water_chem, day_spec_rich)

# interpolated, detrended, etc.
day_spec_rich <- densities.int %>%
  dplyr::filter(species %in% vars, 
                (method=="FlowCam" & density>=100) | (method=="Video" & density>=10) | (method=="Flowcytometer")) %>%
  group_by(bottle, day) %>%
  summarise(day_richness = n()) %>%
  group_by(bottle) %>%
  dplyr::mutate(mean_richness = mean(day_richness, na.rm = T),
                median_richness = median(day_richness, na.rm = T))

dd_list$densities.int.detrend.complete <- full_join(dd_list$densities.int.detrend.complete, day_spec_rich)
dd_list$biomass.int.detrend.complete <- full_join(dd_list$biomass.int.detrend.complete, day_spec_rich)
dd_list$o2.int.detrend.complete <- full_join(dd_list$o2.int.detrend.complete, day_spec_rich)
dd_list$water_chem.int.detrend.complete <- full_join(dd_list$water_chem.int.detrend.complete, day_spec_rich)

dd_list.int$densities.int <- full_join(dd_list.int$densities.int, day_spec_rich)
dd_list.int$biomass.int <- full_join(dd_list.int$biomass.int, day_spec_rich)
dd_list.int$o2.int <- full_join(dd_list.int$o2.int, day_spec_rich)
dd_list.int$water_chem.int <- full_join(dd_list.int$water_chem.int, day_spec_rich)
```


# Saving processed data

```{r}
saveRDS(object = dd_list, file = here("Data/2_TimeSeriesData/Transformed/DensityBiomassTimeSeries.RDS"))
write_csv(o2_bothMeas, here("Data/2_TimeSeriesData/Transformed/OxygenBothSensors.csv"))
saveRDS(object = dd_list.int, file = here("Data/2_TimeSeriesData/Transformed/DensityBiomassTimeSeriesIntermediate.RDS"))
```

