---
title: "Figure 2"
author: "Romana Limberger, Uriah Daugaard et al."
linestretch: 1.5
fontsize: 10pt
output: 
  bookdown::pdf_document2:
    number_sections: true
    toc: true
    toc_depth: 2
    link-citations: true
---

```{r include = FALSE, echo = FALSE}
library(knitr)
opts_chunk$set(
  fig.path='figure/', 
  self.contained=FALSE,
  dev=c('pdf', 'tiff'),
  dpi = 300,
  cache=FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align="center"
)
```

# Packages, data and functions

## Packages

```{r echo = TRUE}
library(tidyverse)  
library(patchwork)  
library(here) 
library(lme4) 
library(lmerTest)
```

## Data
```{r data}
load(here("Data/3_ForecastsResults/3_MergedResults/complete.forecasts.RData"))

complete.forecasts.taxa <- complete.forecasts %>%
  dplyr::filter(
    forecast.method == "Simplex EDM",
    Target != "Total_biomass",
    Target != "Mean_oxygen"
  )
```



## Functions

```{r functionsForConfidenceInterval}
conf_interval <- function(model, nd, response, renameResponse, cmult=1.96){
  nd[,response] <- predict(model, nd, re.form=NA)
  mm <- model.matrix(terms(model),nd)
  pvar1 <- diag(mm %*% tcrossprod(vcov(model),mm))
  nd <- nd %>%
    dplyr::mutate(lower=!!as.name(response)-cmult*sqrt(pvar1),
                  upper=!!as.name(response)+cmult*sqrt(pvar1),
                  Response=renameResponse) %>%
    dplyr::rename(fit=!!as.name(response))
  return(nd)
}
```


# Figure 2

## Models

```{r modelFigure2, message=FALSE, warning=FALSE}
## acf vs exp
m.acf.exp <-
  lmer(
    acf_lag5 ~ median_richness_center * light_treatment + 
      (1 |Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )

## cv vs exp
m.cv.exp <-
  lmer(
    cv.res.log10 ~ median_richness_center * light_treatment + 
      (1 | Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )

### pe vs rmse
m.pe.exp <-
  lmer(
    pe5 ~ median_richness_center * light_treatment + 
      (1 | Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )

## acf vs RMSE
m.acf.rmse <-
  lmer(
    RMSE ~ acf_lag5 * light_treatment + 
      (1 | Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )

## CV vs RMSE
m.cv.rmse <-
  lmer(
    RMSE ~ cv.res.log10 * light_treatment + 
      (1 |Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )

### pe vs rmse
m.pe.rmse <-
  lmer(
    RMSE ~ pe5 * light_treatment + 
      (1 | Target) + (1 | bottle) + (1 | incubator),
    complete.forecasts.taxa
  )
```


## Figure

```{r Figure2Prep, fig.height=7, fig.width=12, message=FALSE, warning=FALSE}
newdat.m.exp <-
  expand.grid(
    median_richness_center = seq(
      min(complete.forecasts.taxa$median_richness_center),
      max(complete.forecasts.taxa$median_richness_center),
      length.out = 100
    ),
    light_treatment = unique(complete.forecasts.taxa$light_treatment)
  )

newdat.m.acf <-
  expand.grid(
    acf_lag5 = seq(
      min(complete.forecasts.taxa$acf_lag5),
      max(complete.forecasts.taxa$acf_lag5),
      length.out = 100
    ),
    light_treatment = unique(complete.forecasts.taxa$light_treatment)
  )

newdat.m.cv <-
  expand.grid(
    cv.res.log10 = seq(min(log10(
      complete.forecasts.taxa$cv.res
    )),
    max(log10(
      complete.forecasts.taxa$cv.res
    )),
    length.out = 100),
    light_treatment = unique(complete.forecasts.taxa$light_treatment)
  )

newdat.m.pe.rmse <-
  expand.grid(
    pe5 = seq(
      min(complete.forecasts.taxa$pe5),
      max(complete.forecasts.taxa$pe5),
      length.out = 100
    ),
    light_treatment = unique(complete.forecasts.taxa$light_treatment)
  )


## acf vs exp
newdat.acf.exp <-
  conf_interval(m.acf.exp, newdat.m.exp, "acf_lag5", "Autocorrelation at lag 5")

## cv vs exp
newdat.cv.exp <-
  conf_interval(m.cv.exp, newdat.m.exp, "cv.res.log10", "cv.res.log10")

## acf vs RMSE
newdat.m.acf.rmse <-
  conf_interval(m.acf.rmse, newdat.m.acf, "RMSE", "Forecast error (RMSE)")

## CV vs RMSE
newdat.m.cv.rmse <-
  conf_interval(m.cv.rmse, newdat.m.cv, "RMSE", "Forecast error (RMSE)")

### pe vs exp design
newdat.m.exp <-
  conf_interval(m.pe.exp, newdat.m.exp, "pe5", "Permutation entropy")

### pe vs rmse
newdat.m.pe.rmse <-
  conf_interval(m.pe.rmse, newdat.m.pe.rmse, "RMSE", "Forecast error (RMSE)")


fig2exp.acf <- complete.forecasts.taxa %>%
  ggplot(aes(median_richness_center, acf_lag5, col = light_treatment)) +
  geom_ribbon(
    data = newdat.acf.exp,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.acf.exp,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(x = "Centered median realized species richness", 
       y = "Autocorrelation at lag 5",
       col = "Light conditions") +
  theme_bw() +
  theme(legend.position = "none")

fig2exp.cv <- complete.forecasts.taxa %>%
  ggplot(aes(median_richness_center, cv.res.log10, col = light_treatment)) +
  geom_ribbon(
    data = newdat.cv.exp,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.cv.exp,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(x = "Centered median realized species richness", 
       y = "log10(Coefficient of variation CV)",
       col = "Light conditions") +
  theme_bw() +
  theme(legend.position = "bottom")

fig2acf.rmse <- complete.forecasts.taxa %>%
  ggplot(aes(acf_lag5, RMSE, col = light_treatment)) +
  geom_ribbon(
    data = newdat.m.acf.rmse,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.m.acf.rmse,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(x = "Autocorrelation at lag 5", y = "Forecast error (RMSE)",
       col = "Light conditions") +
  theme_bw() +
  theme(legend.position = "none")

fig2cv.rmse <- complete.forecasts.taxa %>%
  ggplot(aes(cv.res.log10, RMSE, col = light_treatment)) +
  geom_ribbon(
    data = newdat.m.cv.rmse,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.m.cv.rmse,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(x = "log10(Coefficient of variation CV)",
       y = "Forecast error (RMSE)",
       col = "Light conditions",
       tag = "e") +
  theme_bw() +
  theme(legend.position = "none")

fig2exp.pe <- complete.forecasts.taxa %>%
  dplyr::filter(Target != "Total_biomass",
                Target != "Mean_oxygen",
                forecast.method == "Simplex EDM") %>%
  ggplot(aes(median_richness_center, pe5, col = light_treatment)) +
  geom_ribbon(
    data = newdat.m.exp,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.m.exp,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(x = "Centered median realized species richness", 
       y = "Permutation entropy",
       col = "Light conditions") +
  theme_bw() +
  theme(legend.position = "none")

fig2pe.rmse <- complete.forecasts.taxa %>%
  dplyr::filter(Target != "Total_biomass",
                Target != "Mean_oxygen",
                forecast.method == "Simplex EDM") %>%
  ggplot(aes(pe5, RMSE, col = light_treatment)) +
  geom_ribbon(
    data = newdat.m.pe.rmse,
    aes(
      ymin = lower,
      ymax = upper,
      y = fit,
      group = light_treatment
    ),
    col = NA,
    fill = "grey",
    alpha = 0.3
  ) +
  geom_line(data = newdat.m.pe.rmse,
            size = 1.3,
            aes(y = fit, group = light_treatment)) +
  geom_point(size = 2.5, alpha = 0.6) +
  labs(y = "Forecast error (RMSE)", x = "Permutation entropy",
       col = "Light conditions") +
  theme_bw() +
  theme(legend.position = "none")
```


```{r Figure2, fig.height=7, fig.width=12, message=FALSE, warning=FALSE}
fig2acf.rmse + fig2cv.rmse + fig2pe.rmse +
  fig2exp.acf + fig2exp.cv + fig2exp.pe +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = "a") &
  scale_color_manual(values = c("#C4961A", "#293352"),
                     aesthetics = c("color", "fill")) 
```


**Figure 2**: Relationship of time series metrics with forecast error (a-c), and with experimental treatments (d-f). Time series metrics are (a, d) autocorrelation at lag 5, (b, e) coefficient of variation as a measure of temporal variability, and (c, f) permutation entropy as a measure of time series complexity. Lines display the fit of the corresponding linear mixed model, shaded areas denote the 95% confidence intervals. Taxon, bottle, and incubator were included as random effects in the model. Forecast error is based on forecasts computed with Simplex EDM.

# Versions of R and of the R-packages used


```{r rdetails, echo = FALSE}
# Base packages:
s <- sessionInfo()
s1 <- s$basePkgs[1]

for (i in 2:length(s$basePkgs)){
  s1 <- paste(s1, ", ", s$basePkgs[i], sep = "")
}

# Other Packages:
pack.info <- installed.packages()
output.packages <- data.frame(pack.info[names(s$otherPkgs),
                                        c("Package","Version")])

s2 <- paste(names(s$otherPkgs)[1],
            output.packages[names(s$otherPkgs)[1], "Version"])

k <- length(names(s$otherPkgs))

if(k>1) {
  for (i in 2:k){
    s2 <- paste(s2, ", ",
                paste(names(s$otherPkgs)[i],
                      output.packages[names(s$otherPkgs)[i], "Version"]),
                sep = "")
  }
}  
```


\textbf{\textsf{R} version}: \textsf{`r s$R.version$version.string`}.

\textbf{Base packages}: \textsf{`r s1`}.

\textbf{Other packages}: \textsf{`r s2`}.

\textbf{This document was generated on `r format(Sys.time(), "%B %d, %Y at %H:%M")`}.