---
title: "Forecasting and Calculation of Time Series Metrics"
author: "Romana Limberger, Uriah Daugaard et al."
always_allow_html: true
output:
  html_document:
    number_sections: yes
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
---

<style type="text/css">
.main-container {
max-width: 100% !important;
margin: auto;
}
</style>


# A note

The multiview EDM forecasts are time-costly and can take days. All forecasts are in R-chunks that are not evaluated by default. If you want to run the forecasts change eval=FALSE to eval=TRUE.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies, functions and data

## Dependencies 

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(randomForest)
library(data.table)
library(forecast)
library(parallel)
library(doParallel)
library(rEDM)
library(here)
library(caret)
library(GGally)
library(reticulate) 
library(keras) 

set.seed(1)
options(dplyr.summarise.inform = FALSE)
```

## Functions

```{r load_own_fcts, message=FALSE, warning=FALSE}
source(file = here("R/1_Functions/ForecastingFunctions.R")) ## forecasting functions 
source(file = here("R/1_Functions/PermutationEntropy.R")) ## Permutation entropy function
```


## Data {#data}

```{r dd_list, message=FALSE, warning=FALSE}
dd_list <- readRDS(here("Data/2_TimeSeriesData/Transformed/DensityBiomassTimeSeries.RDS"))
names(dd_list)
```

**Contents of dd_list**

- "biomass" contains the transformed time series of the total biomass 
- "densities" contains the transformed time series of the taxa densities and biomasses
- "o2" contains the transformed oxygen time series
- "water_chem" contains the transformed water chemistry time series (Carbon and Nitrogen)
- "int.detrend.complete" indicates that it contains the respective transformed time series (interpolated, detrended and standardized over the whole duration of the time series)

# Merging of time series data

## Linearly detrended data 

```{r all_data_complete, message=FALSE, warning=FALSE}
complete.ts.dd <- 
  dd_list$biomass.int.detrend.complete %>%
  mutate(variable = "Total biomass",
         value = biomass) %>%
  ungroup() %>%
  dplyr::select(day, bottle, richness, composition, incubator,
                realized_richness, light, light_treatment, 
                light_phase, variable, value, mean_richness, median_richness)

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$densities.int.detrend.complete %>%
              mutate(variable = species,
                     value = density) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$o2.int.detrend.complete %>%
              mutate(variable = "Mean oxygen",
                     value = percent_o2) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$water_chem.int.detrend.complete %>%
              mutate(variable = type,
                     value = concentration) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))


complete.ts.dd <- complete.ts.dd %>%
  pivot_wider(names_from = variable, values_from = value)
```

## Data detrended with segmented approach

```{r all_data_complete_segmented, message=FALSE, warning=FALSE}
complete.ts.dd.seg <- 
  dd_list$biomass.int.detrend.complete %>%
  mutate(variable = "Total biomass",
         value = biomass.seg) %>%
  ungroup() %>%
  dplyr::select(day, bottle, richness, composition, incubator,
                realized_richness, light, light_treatment, 
                light_phase, variable, value, mean_richness, median_richness)

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$densities.int.detrend.complete %>%
              mutate(variable = species,
                     value = density.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$o2.int.detrend.complete %>%
              mutate(variable = "Mean oxygen",
                     value = percent_o2.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$water_chem.int.detrend.complete %>%
              mutate(variable = type,
                     value = concentration.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))


complete.ts.dd.seg <- complete.ts.dd.seg %>%
  pivot_wider(names_from = variable, values_from = value)
```

# Forecasting of abundances, biomass & oxygen


```{r forecast_prep, message=FALSE, warning=FALSE}
additional_targets <- c("Mean oxygen", "Total biomass")

colnames(complete.ts.dd) <- gsub(" ", "_", colnames(complete.ts.dd))
complete.ts.dd.bottle <- split(complete.ts.dd, complete.ts.dd$bottle)

all_possible_predictors <- c(
  "Total_biomass",
  "Bacteria",
  "Chlamydomonas_clumps",
  "Chlamydomonas_reinhardtii",
  "Coleps_sp.",
  "Cosmarium_botrytis",
  "Desmodesmus_armatus",
  "Desmodesmus_clumps",
  "Dividing_Chlamydomonas",
  "Paramecium_bursaria",
  "Paramecium_caudatum",
  "Small_cells",
  "Staurastrum_gracile",
  "Monoraphidium_obtusum",
  "Staurastrum_polytrichum",
  "Stylonychia_sp.",
  "Colpidium_striatum",
  "Dexiostoma_campylum",
  "Euplotes_daidaleos",
  "Loxocephalus_sp.",
  "Mean_oxygen",
  "IC",
  "TC",
  "TN",
  "TOC"
)  
```


## Forecasts of linearly detrended time series

### Multiview EDM

```{r complete_forecasts_mv, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_mvEDM.txt'
file.create(write.file)

# forecasting parallelized for the the bottles.
# with 30 bottles the following number of cores make the most sense: 5, 6, 10, 15, 30
# however, not every bottle will take the same time, because the number of targets and predictors varies...
complete.multiview.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.multiview.forecasts.RDS"))
```


### Simplex EDM

```{r univariatesimplexcomplete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    ddrmse <- data.frame(Target=target, RMSE=rmse, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.simplex.forecasts.RDS"))
```


### ARIMA

```{r arima_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_ARIMA.txt'
file.create(write.file)

complete.arima.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.arima.forecasts.RDS"))
```

### Recurrent neural networks 

```{r RNN_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_RNN.txt'
file.create(write.file)

complete.rnn.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rnn.forecasts.RDS"))
```


### Random forest

```{r RF_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

### Log file
write.file = 'ForecastLog_completeTS_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]
  
  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rf.forecasts.RDS"))
```


## Forecasts of time series detrended in segments

```{r}
colnames(complete.ts.dd.seg) <- gsub(" ", "_", colnames(complete.ts.dd.seg))
complete.ts.dd.seg.bottle <- split(complete.ts.dd.seg, complete.ts.dd.seg$bottle)
```


### Multiview EDM

```{r seg_complete_forecasts_mv, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_mvEDM.txt'
file.create(write.file)

complete.multiview.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.multiview.forecasts.RDS"))
```


### Simplex EDM

```{r seg_univariatesimplexcomplete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    ddrmse <- data.frame(Target=target, RMSE=rmse, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.simplex.forecasts.RDS"))
```


### ARIMA

```{r seg_arima_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_ARIMA.txt'
file.create(write.file)

complete.arima.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.arima.forecasts.RDS"))
```

### Recurrent neural networks

```{r seg_RNN_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_RNN.txt'
file.create(write.file)

complete.rnn.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    # install_keras()
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rnn.forecasts.RDS"))
```


### Random forest

```{r seg_RF_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

### Log file
write.file = 'seg_ForecastLog_completeTS_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]
  
  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rf.forecasts.RDS"))
```



## In-sample based detrending and subsequent forecasting

### Data preparation

#### In sample processing

```{r in_sample_ts_processing, message=FALSE, warning=FALSE}
# load in intermediate data (only interpolated)
dd_list.int <- readRDS(here("Data/2_TimeSeriesData/Transformed/DensityBiomassTimeSeriesIntermediate.RDS"))

##############################################################
## remove linear trend

predict_lm_indexed <- function(var,day,index){
  if(sum(is.na(var))==length(var)){
    return(var)
  }
  var2 <- var[index]
  day2 <- day[index]
  m <- lm(var2~day2)
  predicted_residuals <- var - predict(m, newdata=data.frame(day2=day,var2=var)) 
  return(predicted_residuals)
}

ind <- 1:111


## total biomass

biomass.insample.detrended <- setDT(dd_list.int$biomass.int)[, list(day = day,
                                                                    biomass = predict_lm_indexed(biomass,day,ind)),
                                                             by = .(bottle,richness,
                                                                    composition,incubator,
                                                                    light_treatment, realized_richness)]


## species densities and biomass

densities.insample.detrended <- dd_list.int$densities.int %>%
  dplyr::filter(!(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02"))

densities.insample.detrended <- 
  setDT(densities.insample.detrended)[, list(day = day,
                                             density = predict_lm_indexed(density,day,ind),
                                             biomass = predict_lm_indexed(biomass,day,ind)), 
                                      by = .(species,bottle,richness,
                                             composition,incubator,
                                             method,light_treatment,ForecastTarget,
                                             realized_richness)]

DesmoB2.insample.detrended <- dd_list.int$densities.int %>%
  dplyr::filter(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02")

DesmoB2.insample.detrended <-
  setDT(DesmoB2.insample.detrended)[, list(day = day,
                                           density = predict_lm_indexed(density,day,ind),
                                           biomass = predict_lm_indexed(biomass,day,ind)), 
                                    by = .(species,bottle,richness,
                                           composition,incubator,
                                           method,light_treatment,ForecastTarget,
                                           realized_richness)]

## o2

o2.insample.detrended <- setDT(dd_list.int$o2.int)[, list(day = day,
                                                          percent_o2 = predict_lm_indexed(percent_o2,day,ind)),
                                                   by = .(bottle,
                                                          composition, richness,incubator,
                                                          method,light_treatment, realized_richness)]



## water chem

chem.insample.detrended <- 
  setDT(dd_list.int$water_chem.int)[, list(day = day,
                                           concentration = predict_lm_indexed(concentration,day,ind)),
                                    by = .(type,bottle,
                                           composition, richness,incubator,
                                           light_treatment, realized_richness)]


##############################################################
## Standardization

biomass.insample.detrended <- biomass.insample.detrended %>%
  group_by(bottle, richness, light_treatment, composition, incubator, realized_richness) %>%
  mutate(biomass = biomass/sd(biomass[ind], na.rm = T))

densities.insample.detrended <- densities.insample.detrended %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  mutate(density = density/sd(density[ind], na.rm = T), 
         biomass = biomass/sd(biomass[ind], na.rm = T),
         density = ifelse(is.nan(density),0,density), # because didinium
         biomass = ifelse(is.nan(biomass),0,biomass))

DesmoB2.insample.detrended <- DesmoB2.insample.detrended %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  mutate(density = density/sd(density[ind], na.rm = T), 
         biomass = biomass/sd(biomass[ind], na.rm = T)) 

densities.insample.detrended <- rbind(densities.insample.detrended,
                                      DesmoB2.insample.detrended)

o2.insample.detrended <- o2.insample.detrended %>%
  group_by(bottle, light_treatment, composition, incubator, method, richness, realized_richness) %>%
  mutate(percent_o2 = percent_o2/sd(percent_o2[ind], na.rm = T))

chem.insample.detrended <- chem.insample.detrended %>%
  group_by(bottle, light_treatment, composition, incubator, type, richness, realized_richness) %>%
  mutate(concentration = concentration/sd(concentration, na.rm = T))

```

#### Merging data

```{r all_data_complete_insample, message=FALSE, warning=FALSE}
vars <- c("Coleps sp.","Paramecium bursaria","Paramecium caudatum", 
          "Stylonychia sp.","Colpidium striatum","Euplotes daidaleos",
          "Dexiostoma campylum","Loxocephalus sp.","Chlamydomonas reinhardtii",
          "Cosmarium botrytis",
          "Desmodesmus armatus", "Staurastrum gracile","Monoraphidium obtusum",
          "Staurastrum polytrichum","Bacteria") 

day_spec_rich <- dd_list.int$densities.int %>%
  dplyr::filter(species %in% vars, 
                (method=="FlowCam" & density>=100) | (method=="Video" & density>=10) | (method=="Flowcytometer")) %>%
  group_by(bottle, day) %>%
  summarise(day_richness = n()) %>%
  group_by(bottle) %>%
  dplyr::mutate(mean_richness = mean(day_richness, na.rm = T),
                median_richness = median(day_richness, na.rm = T))

biomass.insample.detrended <- full_join(biomass.insample.detrended, day_spec_rich)
densities.insample.detrended <- full_join(densities.insample.detrended, day_spec_rich)
o2.insample.detrended <- full_join(o2.insample.detrended, day_spec_rich)
chem.insample.detrended <- full_join(chem.insample.detrended, day_spec_rich)


in.sample.complete <- 
  biomass.insample.detrended %>%
  mutate(variable = "Total biomass",
         value = biomass) %>%
  ungroup() %>%
  dplyr::select(day, bottle, richness, composition, incubator,
                realized_richness, light_treatment, 
                variable, value, mean_richness, median_richness)

in.sample.complete <-
  full_join(in.sample.complete,
            densities.insample.detrended %>%
              mutate(variable = species,
                     value = density) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light_treatment, 
                            variable, value, mean_richness, median_richness))

in.sample.complete <-
  full_join(in.sample.complete,
            o2.insample.detrended %>%
              mutate(variable = "Mean oxygen",
                     value = percent_o2) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light_treatment, 
                            variable, value, mean_richness, median_richness))

in.sample.complete <-
  full_join(in.sample.complete,
            chem.insample.detrended %>%
              mutate(variable = type,
                     value = concentration) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light_treatment, 
                            variable, value, mean_richness, median_richness))


in.sample.complete <- in.sample.complete %>%
  pivot_wider(names_from = variable, values_from = value)

in.sample.complete <- in.sample.complete %>%
  dplyr::mutate(light = case_when(light_treatment =="constant" ~ 30,
                                  day >= 191 ~ 1,
                                  184 <= day & day < 191 ~ 2,
                                  177 <= day & day < 184 ~ 4,
                                  170 <= day & day < 177 ~ 6,
                                  163 <= day & day < 170 ~ 8,
                                  156 <= day & day < 163 ~ 10,
                                  149 <= day & day < 156 ~ 12,
                                  142 <= day & day < 149 ~ 14,
                                  135 <= day & day < 142 ~ 16,
                                  128 <= day & day < 135 ~ 18,
                                  121 <= day & day < 128 ~ 20,
                                  114 <= day & day < 121 ~ 22,
                                  107 <= day & day < 114 ~ 24,
                                  100 <= day & day < 107 ~ 26,
                                  93 <= day & day < 100 ~ 28,
                                  day < 93 ~ 30),
                light_phase = case_when(light_treatment =="constant" ~ "constant",
                                        day < 93 ~ "constant before decrease",
                                        day >= 93 & day < 198 ~ "decreasing",
                                        day >= 198 ~ "constant after decrease"))
```


```{r forecast_prep_insample, message=FALSE, warning=FALSE}
colnames(in.sample.complete) <- gsub(" ", "_", colnames(in.sample.complete))
in.sample.complete.bottle <- split(in.sample.complete, in.sample.complete$bottle)

dir.create(paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults"))
```


### Forecasting based on in-sample processing

#### Multiview EDM

```{r complete_forecasts_mv_inSample, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
# targets <- dd_list$densities.int.detrend.complete %>% dplyr::filter(ForecastTarget==T)
# targets <- unique(targets$species) # targets should be: bact, chlamy clumps, chlamy, coleps, desmo, desmo clumps, dividing chlamy, caudatum, bursaria, small cells, mono, stylo,  dexio, stau gracile

set.seed(1231)

### Log file
write.file = 'ForecastLog_inSample_mvEDM.txt'
file.create(write.file)

# forecasting parallelized for the the bottles.
# with 30 bottles the following number of cores make the most sense: 5, 6, 10, 15, 30
# however, not every bottle will take the same time, because the number of targets and predictors varies...
complete.multiview.forecasts <- mclapply(in.sample.complete.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  # change length.out until desired length is achieved after removing duplicates. 25?
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  
  # ##### for testing (to delete)
  # ks <- 3
  # possible_predictors <- possible_predictors[1:5]
  # targets <- targets[1:5]
  
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts,
        file = paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults/complete.multiview.forecasts.RDS"))
```


#### Simplex EDM

```{r univariatesimplexcomplete_inSample, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_inSample_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(in.sample.complete.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  
  # ##### for testing (to delete)
  # targets <- targets[1]
  
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    ddrmse <- data.frame(Target=target, RMSE=rmse, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults/complete.simplex.forecasts.RDS"))
```


#### ARIMA

```{r arima_forecast_complete_inSample, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_inSample_ARIMA.txt'
file.create(write.file)

complete.arima.forecasts <- mclapply(in.sample.complete.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  # ##### for testing (to delete)
  # covariates <- covariates[1:5]
  # targets <- targets[1:5]
  
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults/complete.arima.forecasts.RDS"))
```

#### Recurrent neural networks 

```{r RNN_forecast_complete_inSample, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

set.seed(1231)

### Log file
write.file = 'ForecastLog_inSample_RNN.txt'
file.create(write.file)

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}


complete.rnn.forecasts <- mclapply(in.sample.complete.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  # ##### for testing (to delete)
  # possible_predictors <- possible_predictors[1:5]
  # targets <- targets[1:5]
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    # install_keras()
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=cr)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults/complete.rnn.forecasts.RDS"))
```


#### Random forest

```{r RF_forecast_complete_inSample, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

### Log file
write.file = 'ForecastLog_inSample_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(in.sample.complete.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]
  
  # ##### for testing (to delete)
  # possible_predictors <- possible_predictors[1:5]
  # targets <- targets[1:5]
  
  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/5_InSampleProcessResults/complete.rf.forecasts.RDS"))
```


## Rerun of forecasts for the back-transformation before the RMSE calculation

### Data preparation

```{r SdMeanData, message=FALSE, warning=FALSE}
## total biomass
biomassMeanSd <- setDT(dd_list.int$biomass.int)[, list(TimeSeriesSD = sd(biomass, na.rm=T),
                                                       TimeSeriesSDLast = sd(tail(biomass,12), na.rm = T),
                                                       RemovedTrend = coef(lm(biomass~day))[2]),
                                                by = .(bottle,richness,
                                                       composition,incubator,
                                                       light_treatment, realized_richness)]

biomassMeanSd2 <- setDT(dd_list.int$biomass.int)[, list(day = day,
                                                        biomass = residuals(lm(biomass~day))),
                                                 by = .(bottle,richness,
                                                        composition,incubator,
                                                        light_treatment, realized_richness)]

## species densities and biomass

densityMeanSd <- dd_list.int$densities.int %>%
  filter(!(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02"))
densityMeanSd <- 
  setDT(densityMeanSd)[, list(TimeSeriesSD = sd(density, na.rm=T),
                              TimeSeriesSDLast = sd(tail(density,12), na.rm = T),
                              RemovedTrend = coef(lm(density~day))[2]), # density detrended linear simple
                       by = .(species,bottle,richness,
                              composition,incubator,
                              method,light_treatment,ForecastTarget,
                              realized_richness)]

B2MeanSd <- dd_list.int$densities.int %>%
  filter(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02")
B2MeanSd <- setDT(B2MeanSd)[, list(TimeSeriesSD = sd(density, na.rm=T),
                                   TimeSeriesSDLast = sd(tail(density,12), na.rm = T),
                                   RemovedTrend = coef(lm(density~day))[2]),
                            by = .(species,bottle,richness,
                                   composition,incubator,
                                   method,light_treatment,ForecastTarget,
                                   realized_richness)]

densityMeanSd <- rbind(densityMeanSd,
                       B2MeanSd)


densityMeanSd2 <- dd_list.int$densities.int %>%
  dplyr::filter(!(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02"))
densityMeanSd2 <- 
  setDT(densityMeanSd2)[, list(day = day,
                               density = residuals(lm(density~day))), 
                        by = .(species,bottle,richness,
                               composition,incubator,
                               method,light_treatment,ForecastTarget,
                               realized_richness)]

B2MeanSd2 <- dd_list.int$densities.int %>%
  dplyr::filter(species %in% c("Desmodesmus armatus","Desmodesmus clumps") & bottle=="b_02")
B2MeanSd2 <-
  setDT(B2MeanSd2)[, list(day = day,
                          density = residuals(lm(density~day))), 
                   by = .(species,bottle,richness,
                          composition,incubator,
                          method,light_treatment,ForecastTarget,
                          realized_richness)]


## o2

o2MeanSd <- setDT(dd_list.int$o2.int)[, list(TimeSeriesSD = sd(percent_o2, na.rm=T),
                                             TimeSeriesSDLast = sd(tail(percent_o2,12), na.rm = T),
                                             RemovedTrend = coef(lm(percent_o2~day))[2]),
                                      by = .(bottle,
                                             composition, richness,incubator,
                                             method,light_treatment, realized_richness)]

o2MeanSd2 <- setDT(dd_list.int$o2.int)[, list(day = day,
                                              percent_o2 = residuals(lm(percent_o2~day))),
                                       by = .(bottle,
                                              composition, richness,incubator,
                                              method,light_treatment, realized_richness)]

## water chem

chemMeanSd <- setDT(dd_list.int$water_chem.int)[, list(TimeSeriesSD = sd(concentration, na.rm=T),
                                                       TimeSeriesSDLast = sd(tail(concentration,12), na.rm = T),
                                                       RemovedTrend = coef(lm(concentration~day))[2]),
                                                by = .(type,bottle,
                                                       composition, richness,incubator,
                                                       light_treatment, realized_richness)]

chemMeanSd2 <- 
  setDT(dd_list.int$water_chem.int)[, list(day = day,
                                           concentration = residuals(lm(concentration~day))),
                                    by = .(type,bottle,
                                           composition, richness,incubator,
                                           light_treatment, realized_richness)]

##############################################################
## Standardization

biomassMeanSd2 <- biomassMeanSd2 %>%
  group_by(bottle, richness, light_treatment, composition, incubator, realized_richness) %>%
  summarize(ResidualSD = sd(biomass, na.rm = T))

densityMeanSd2 <- densityMeanSd2 %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  summarize(ResidualSD = sd(density, na.rm = T)) 

B2MeanSd2 <- B2MeanSd2 %>%
  group_by(species, bottle, richness, light_treatment, composition, incubator, method, realized_richness) %>%
  summarize(ResidualSD = sd(density, na.rm = T)) 

densityMeanSd2 <- rbind(densityMeanSd2,
                        B2MeanSd2)

o2MeanSd2 <- o2MeanSd2 %>%
  group_by(bottle, light_treatment, composition, incubator, method, richness, realized_richness) %>%
  summarize(ResidualSD = sd(percent_o2, na.rm = T)) 

chemMeanSd2 <- chemMeanSd2 %>%
  group_by(bottle, light_treatment, composition, incubator, type, richness, realized_richness) %>%
  summarize(ResidualSD = sd(concentration, na.rm = T)) 

### merge

biomassMeanSd <- full_join(biomassMeanSd, biomassMeanSd2) %>%
  dplyr::mutate(Target = "Total_biomass")
densityMeanSd <- full_join(densityMeanSd, densityMeanSd2) %>%
  dplyr::rename(Target = species)
o2MeanSd <- full_join(o2MeanSd, o2MeanSd2) %>%
  dplyr::mutate(Target = "Mean_oxygen")
chemMeanSd <- full_join(chemMeanSd, chemMeanSd2) %>%
  dplyr::rename(Target = type)

MeansSD <- full_join(biomassMeanSd,
                     full_join(densityMeanSd,
                               full_join(o2MeanSd, chemMeanSd))) %>%
  dplyr::mutate(Target = gsub(" ", "_", Target))
```

```{r forecast_prep_rmseDirectory, message=FALSE, warning=FALSE}
dir.create(paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE"))
```


### Multiview EDM

```{r complete_forecasts_mv_backTransf, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
# targets <- dd_list$densities.int.detrend.complete %>% dplyr::filter(ForecastTarget==T)
# targets <- unique(targets$species) # targets should be: bact, chlamy clumps, chlamy, coleps, desmo, desmo clumps, dividing chlamy, caudatum, bursaria, small cells, mono, stylo,  dexio, stau gracile

set.seed(1231)

### Log file
write.file = 'ForecastLog_RMSEBackT_mvEDM.txt'
file.create(write.file)

# forecasting parallelized for the the bottles.
# with 30 bottles the following number of cores make the most sense: 5, 6, 10, 15, 30
# however, not every bottle will take the same time, because the number of targets and predictors varies...
complete.multiview.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  # change length.out until desired length is achieved after removing duplicates. 25?
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  
  # ##### for testing (to delete)
  # ks <- 1:3
  # possible_predictors <- possible_predictors[1:3]
  # targets <- targets[1:3]
  
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file, 
                             backTransform = T, MeansSD = MeansSD)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts,
        file = paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE/complete.multiview.forecasts.RDS"))
```


### Simplex EDM

```{r univariatesimplexcomplete_backTransf, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_RMSEBackT_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  
  # ##### for testing (to delete)
  # targets <- targets[1]
  
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    
    MeanSdFiltered <- MeansSD %>%
      dplyr::filter(bottle == unique(bottle_specs$bottle),
                    Target == target)
    
    backTrend <- unique(MeanSdFiltered$RemovedTrend)
    backSD <- unique(MeanSdFiltered$ResidualSD)
    sdTS <- unique(MeanSdFiltered$TimeSeriesSD)
    
    forecast <- forecast %>% na.omit()
    obs <- forecast$Observations*backSD + backTrend*forecast$day
    preds <- forecast$Predictions*backSD + backTrend*forecast$day
    rmseBackT <- sqrt(mean((obs-preds)^2, na.rm = T))/sdTS
    
    ddrmse <- data.frame(Target=target, RMSE=rmse, RMSE_BackT = rmseBackT, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE/complete.simplex.forecasts.RDS"))
```


### ARIMA

```{r arima_forecast_complete_backTransf, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_RMSEBackT_ARIMA.txt'
file.create(write.file)

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}

complete.arima.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  # ##### for testing (to delete)
  # covariates <- covariates[1:5]
  # targets <- targets[1:5]
  
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    MeanSdFiltered <- MeansSD %>%
      dplyr::filter(bottle == unique(bottle_specs$bottle),
                    Target == target)
    
    backTrend <- unique(MeanSdFiltered$RemovedTrend)
    backSD <- unique(MeanSdFiltered$ResidualSD)
    sdTS <- unique(MeanSdFiltered$TimeSeriesSD)
    
    days <- 112:123*(7/3)
    obs <- unlist(b.forecast[target]*backSD + backTrend*days)
    preds <- forecast*backSD + backTrend*days
    rmseBackT <- sqrt(mean((obs-preds)^2, na.rm = T))/sdTS
    
    ddrmse <- data.frame(Target=target, RMSE=rmse, RMSE_BackT = rmseBackT)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = cr) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE/complete.arima.forecasts.RDS"))
```

### Recurrent neural networks 

```{r RNN_forecast_complete_backTransf, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}

set.seed(1231)

### Log file
write.file = 'ForecastLog_RMSEBackT_RNN.txt'
file.create(write.file)

complete.rnn.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  # ##### for testing (to delete)
  # possible_predictors <- possible_predictors[1:5]
  # targets <- targets[1:5]
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    # install_keras()
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    MeanSdFiltered <- MeansSD %>%
      dplyr::filter(bottle == unique(bottle_specs$bottle),
                    Target == target)
    
    backTrend <- unique(MeanSdFiltered$RemovedTrend)
    backSD <- unique(MeanSdFiltered$ResidualSD)
    sdTS <- unique(MeanSdFiltered$TimeSeriesSD)
    
    days <- 113:123*(7/3)
    obs <- unlist(y_test*backSD + backTrend*days)
    preds <- unlist(forecast*backSD + backTrend*days)
    rmseBackT <- sqrt(mean((obs-preds)^2, na.rm = T))/sdTS
    
    ddrmse <- data.frame(Target=target, RMSE=rmse, RMSE_BackT = rmseBackT)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=cr)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE/complete.rnn.forecasts.RDS"))
```


### Random forest

```{r RF_forecast_complete_backTransf, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

if(.Platform$OS.type=="windows"){
  cr <- 1
} else {
  cr <- detectCores()-2
}

### Log file
write.file = 'ForecastLog_completeTS_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  # complete ts: forecast the last 12 time steps (ca. 10%)? (i.e last 4 weeks)
  # phase ts: forecast the last 9 time steps (ca. 25%)? (i.e. last 3 weeks of a phase)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]
  
  # ##### for testing (to delete)
  # possible_predictors <- possible_predictors[1:5]
  # targets <- targets[1:5]
  
  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    MeanSdFiltered <- MeansSD %>%
      dplyr::filter(bottle == unique(bottle_specs$bottle),
                    Target == target)
    
    backTrend <- unique(MeanSdFiltered$RemovedTrend)
    backSD <- unique(MeanSdFiltered$ResidualSD)
    sdTS <- unique(MeanSdFiltered$TimeSeriesSD)
    
    days <- 112:123*(7/3)
    obs <- unlist(forecasts_rf$observed*backSD + backTrend*days)
    preds <- unlist(forecasts_rf$predicted*backSD + backTrend*days)
    rmseBackT <- sqrt(mean((obs-preds)^2, na.rm = T))/sdTS
    
    ddrmse <- data.frame(Target=target, RMSE=rmse, RMSE_BackT = rmseBackT)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=cr)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = paste0(here(),"/Data/3_ForecastsResults/4_BackTransformedRMSE/complete.rf.forecasts.RDS"))
```





## Post-forecasting data wrangling 

```{r forecast_post_processing, message=FALSE, warning=FALSE}
# back-transformed forecasts for RMSE

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/4_BackTransformedRMSE/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/4_BackTransformedRMSE/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/4_BackTransformedRMSE/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/4_BackTransformedRMSE/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/4_BackTransformedRMSE/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"

RMSEBackT.complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                                   complete.multiview.forecasts,
                                                   complete.rnn.forecasts,
                                                   complete.simplex.forecasts,
                                                   complete.rf.forecasts), 
                                              dplyr::full_join)

# based on in sample detrending

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/5_InSampleProcessResults/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/5_InSampleProcessResults/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/5_InSampleProcessResults/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/5_InSampleProcessResults/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/5_InSampleProcessResults/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"

inSample.complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                                  complete.multiview.forecasts,
                                                  complete.rnn.forecasts,
                                                  complete.simplex.forecasts,
                                                  complete.rf.forecasts), 
                                             dplyr::full_join)

# based on segmented regression detrending

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"


seg.complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                             complete.multiview.forecasts,
                                             complete.rnn.forecasts,
                                             complete.simplex.forecasts,
                                             complete.rf.forecasts), 
                                        dplyr::full_join)

# based on simple temporal regression detrending

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"

complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                         complete.multiview.forecasts,
                                         complete.rnn.forecasts,
                                         complete.rf.forecasts,
                                         complete.simplex.forecasts), 
                                    dplyr::full_join)
```


# Calculation of temporal stability and permutation entropy

```{r complete.CV.calc, message=FALSE, warning=FALSE, cache=FALSE}
biomass.complete.pe.cv <- dd_list$biomass.int.detrend.complete %>%
  group_by(bottle, composition, richness, incubator, light_treatment, realized_richness, mean_richness, median_richness) %>%
  summarize(biomass.pe.5 = PE(x = biomass, weighted = T, word_length = 5, tau = 1,
                              tie_method = "noise", noise_amount = 10),
            biomass.cv.res = sd(biomass.detrended.unstandardized, na.rm=T)/mean(biomass.original, na.rm=T),
            biomass.acf_lag5 = stats::acf(biomass, lag.max=5, plot=F, na.action = na.pass)$acf[6])

density.complete.pe.cv <- dd_list$densities.int.detrend.complete %>%
  group_by(species, bottle, composition, ForecastTarget, richness, incubator, method, light_treatment, realized_richness, mean_richness, median_richness) %>%
  summarize(density.pe.5 = PE(x = density, weighted = T, word_length = 5, tau = 1,
                              tie_method = "noise", noise_amount = 10),
            biomass.pe.5 = ifelse(all(is.na(biomass)), NA, PE(x = biomass, weighted = T, word_length = 5, tau = 1,
                                                              tie_method = "noise", noise_amount = 10)),
            biomass.cv.res = sd(biomass.detrended.unstandardized, na.rm=T)/mean(biomass.original, na.rm=T),
            density.cv.res = sd(density.detrended.unstandardized, na.rm=T)/mean(density.original, na.rm=T),
            biomass.acf_lag5 = stats::acf(biomass, lag.max=5, plot=F, na.action = na.pass)$acf[6],
            density.acf_lag5 = stats::acf(density, lag.max=5, plot=F, na.action = na.pass)$acf[6])

o2.complete.pe.cv <- dd_list$o2.int.detrend.complete %>%
  group_by(bottle, composition, light_treatment, richness, realized_richness, mean_richness, median_richness) %>%
  summarize(percent_o2.pe.5 = PE(x = percent_o2, weighted = T, word_length = 5, tau = 1,
                                 tie_method = "noise", noise_amount = 10),
            percent_o2.cv.res = sd(percent_o2.detrended.unstandardized, na.rm=T)/mean(percent_o2.original, na.rm=T),
            percent_o2.acf_lag5 = stats::acf(percent_o2, lag.max=5, plot=F, na.action = na.pass)$acf[6])

water_chem.complete.pe.cv <- dd_list$water_chem.int.detrend.complete %>%
  group_by(type, bottle, composition, light_treatment, richness, realized_richness, mean_richness, median_richness) %>%
  summarize(concentration.pe.5 = PE(x = concentration, weighted = T, word_length = 5, tau = 1,
                                    tie_method = "noise", noise_amount = 10),
            concentration.cv.res = sd(concentration.detrended.unstandardized, na.rm=T)/mean(concentration.original, na.rm=T),
            concentration.acf_lag5 = stats::acf(concentration, lag.max=5, plot=F, na.action = na.pass)$acf[6])
```


## Addition to forecast results

```{r mergeForecastsandTSMetrics, warning=FALSE, message=FALSE}
complete.forecasts <- complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))

seg.complete.forecasts <- seg.complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))

inSample.complete.forecasts <- inSample.complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))

RMSEBackT.complete.forecasts <- RMSEBackT.complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))

complete.forecasts.pe.cv <- full_join(complete.forecasts %>%
                                        dplyr::filter(Target == "Total_biomass"),
                                      biomass.complete.pe.cv %>%
                                        mutate(Target = "Total_biomass") %>%
                                        rename(cv.res=biomass.cv.res,
                                               acf_lag5 = biomass.acf_lag5,
                                               pe5=biomass.pe.5))

complete.forecasts.pe.cv2 <- full_join(complete.forecasts %>%
                                         dplyr::filter(Target != "Total_biomass",Target != "Mean_oxygen"),
                                       density.complete.pe.cv %>%
                                         ungroup() %>%
                                         mutate(Target=gsub(" ", "_", species)) %>%
                                         dplyr::select(-biomass.pe.5, -biomass.cv.res, -biomass.acf_lag5,
                                                       -species, -method, -ForecastTarget) %>%
                                         rename(cv.res=density.cv.res,
                                                acf_lag5 = density.acf_lag5,
                                                pe5=density.pe.5)) %>%
  drop_na(RMSE)

complete.forecasts.pe.cv3 <- full_join(complete.forecasts %>%
                                         dplyr::filter(Target == "Mean_oxygen"),
                                       o2.complete.pe.cv %>%
                                         mutate(Target = "Mean_oxygen") %>%
                                         rename(cv.res=percent_o2.cv.res,
                                                acf_lag5 = percent_o2.acf_lag5,
                                                pe5=percent_o2.pe.5))

complete.forecasts.pe.cv <- rbind(complete.forecasts.pe.cv,
                                  complete.forecasts.pe.cv2,
                                  complete.forecasts.pe.cv3)

complete.forecasts <- complete.forecasts.pe.cv %>%
  dplyr::select("bottle","richness","composition","incubator","realized_richness",
                "light_treatment","Target", "RMSE","forecast.method",
                "mean_richness","median_richness","realized_richness_center",
                "median_richness_center","mean_richness_center","pe5",
                "cv.res","acf_lag5") %>%
  mutate(cv.res.log10 = log10(cv.res))

seg.complete.forecasts <- seg.complete.forecasts %>%
  dplyr::select("bottle","richness","composition","incubator","realized_richness",
                "light_treatment","Target", "RMSE","forecast.method",
                "mean_richness","median_richness","realized_richness_center",
                "median_richness_center","mean_richness_center")
```


# Saving merged and processed results

```{r}
save(complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/complete.forecasts.RData"))
save(seg.complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/seg.complete.forecasts.RData"))
save(RMSEBackT.complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/RMSEBackT.complete.forecasts.RData"))
save(inSample.complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/inSample.complete.forecasts.RData"))
``` 

