---
title: "Forecasting and Calculation of Time Series Metrics"
author: "Romana Limberger, Uriah Daugaard et al."
always_allow_html: true
output:
  html_document:
    number_sections: yes
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
---

<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
</style>


# A note

The multiview EDM forecasts are time-costly and can take days. All forecasts are in R-chunks that are not evaluated by default. If you want to run the forecasts change eval=FALSE to eval=TRUE.
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies, functions and data

## Dependencies 

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(randomForest)
library(data.table)
library(forecast)
library(parallel)
library(doParallel)
library(rEDM)
library(here)
library(caret)
library(GGally)
library(reticulate) 
library(keras) 

set.seed(1)
options(dplyr.summarise.inform = FALSE)
```

## Functions

```{r load_own_fcts, message=FALSE, warning=FALSE}
source(file = here("R/1_Functions/ForecastingFunctions.R")) ## forecasting functions 
source(file = here("R/1_Functions/PermutationEntropy.R")) ## Permutation entropy function
```


## Data {#data}

```{r dd_list, message=FALSE, warning=FALSE}
dd_list <- readRDS(here("Data/2_TimeSeriesData/Transformed/DensityBiomassTimeSeries.RDS"))
names(dd_list)
```

**Contents of dd_list**

- "biomass" contains the transformed time series of the total biomass 
- "densities" contains the transformed time series of the taxa densities and biomasses
- "o2" contains the transformed oxygen time series
- "water_chem" contains the transformed water chemistry time series (Carbon and Nitrogen)
- "int.detrend.complete" indicates that it contains the respective transformed time series (interpolated, detrended and standardized over the whole duration of the time series)

# Merging of time series data

## Linearly detrended data 

```{r all_data_complete, message=FALSE, warning=FALSE}
complete.ts.dd <- 
  dd_list$biomass.int.detrend.complete %>%
  mutate(variable = "Total biomass",
         value = biomass) %>%
  ungroup() %>%
  dplyr::select(day, bottle, richness, composition, incubator,
                realized_richness, light, light_treatment, 
                light_phase, variable, value, mean_richness, median_richness)

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$densities.int.detrend.complete %>%
              mutate(variable = species,
                     value = density) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$o2.int.detrend.complete %>%
              mutate(variable = "Mean oxygen",
                     value = percent_o2) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd <-
  full_join(complete.ts.dd,
            dd_list$water_chem.int.detrend.complete %>%
              mutate(variable = type,
                     value = concentration) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))


complete.ts.dd <- complete.ts.dd %>%
  pivot_wider(names_from = variable, values_from = value)
```

## Data detrended with segmented approach

```{r all_data_complete_segmented, message=FALSE, warning=FALSE}
complete.ts.dd.seg <- 
  dd_list$biomass.int.detrend.complete %>%
  mutate(variable = "Total biomass",
         value = biomass.seg) %>%
  ungroup() %>%
  dplyr::select(day, bottle, richness, composition, incubator,
                realized_richness, light, light_treatment, 
                light_phase, variable, value, mean_richness, median_richness)

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$densities.int.detrend.complete %>%
              mutate(variable = species,
                     value = density.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$o2.int.detrend.complete %>%
              mutate(variable = "Mean oxygen",
                     value = percent_o2.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))

complete.ts.dd.seg <-
  full_join(complete.ts.dd.seg,
            dd_list$water_chem.int.detrend.complete %>%
              mutate(variable = type,
                     value = concentration.seg) %>%
              ungroup() %>%
              dplyr::select(day, bottle, richness, composition, incubator,
                            realized_richness, light, light_treatment, 
                            light_phase, variable, value, mean_richness, median_richness))


complete.ts.dd.seg <- complete.ts.dd.seg %>%
  pivot_wider(names_from = variable, values_from = value)
```

# Forecasting of abundances, biomass & oxygen


```{r forecast_prep, message=FALSE, warning=FALSE}
additional_targets <- c("Mean oxygen", "Total biomass")

colnames(complete.ts.dd) <- gsub(" ", "_", colnames(complete.ts.dd))
complete.ts.dd.bottle <- split(complete.ts.dd, complete.ts.dd$bottle)

all_possible_predictors <- c(
  "Total_biomass",
  "Bacteria",
  "Chlamydomonas_clumps",
  "Chlamydomonas_reinhardtii",
  "Coleps_sp.",
  "Cosmarium_botrytis",
  "Desmodesmus_armatus",
  "Desmodesmus_clumps",
  "Dividing_Chlamydomonas",
  "Paramecium_bursaria",
  "Paramecium_caudatum",
  "Small_cells",
  "Staurastrum_gracile",
  "Monoraphidium_obtusum",
  "Staurastrum_polytrichum",
  "Stylonychia_sp.",
  "Colpidium_striatum",
  "Dexiostoma_campylum",
  "Euplotes_daidaleos",
  "Loxocephalus_sp.",
  "Mean_oxygen",
  "IC",
  "TC",
  "TN",
  "TOC"
)  
```


## Forecasts of linearly detrended time series

### Multiview EDM

```{r complete_forecasts_mv, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_mvEDM.txt'
file.create(write.file)

# forecasting parallelized for the the bottles.
# with 30 bottles the following number of cores make the most sense: 5, 6, 10, 15, 30
# however, not every bottle will take the same time, because the number of targets and predictors varies...
complete.multiview.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.multiview.forecasts.RDS"))
```


### Simplex EDM

```{r univariatesimplexcomplete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    ddrmse <- data.frame(Target=target, RMSE=rmse, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.simplex.forecasts.RDS"))
```


### ARIMA

```{r arima_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_ARIMA.txt'
file.create(write.file)

complete.arima.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.arima.forecasts.RDS"))
```

### Recurrent neural networks 

```{r RNN_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

set.seed(1231)

### Log file
write.file = 'ForecastLog_completeTS_RNN.txt'
file.create(write.file)

complete.rnn.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rnn.forecasts.RDS"))
```


### Random forest

```{r RF_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

### Log file
write.file = 'ForecastLog_completeTS_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(complete.ts.dd.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]

  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rf.forecasts.RDS"))
```


## Forecasts of time series detrended in segments

```{r}
colnames(complete.ts.dd.seg) <- gsub(" ", "_", colnames(complete.ts.dd.seg))
complete.ts.dd.seg.bottle <- split(complete.ts.dd.seg, complete.ts.dd.seg$bottle)
```


### Multiview EDM

```{r seg_complete_forecasts_mv, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_mvEDM.txt'
file.create(write.file)

complete.multiview.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  ks <- unique(round(exp(seq(log(0.8),log(ceiling(percent*nrow(dd.bottle))),length.out = 42))))
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  ### Forecasting
  df <- model_fitter_wrapper(whole.data = dd.bottle,
                             predictor_combinations = list(possible_predictors),
                             targets = targets,
                             lib = lib, pred = pred,
                             num.clusters = 1,
                             E = 3, max_lag = 4, k=ks, Tp = 1,
                             predictors = possible_predictors, 
                             write.file = write.file)
  
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.multiview.forecasts <- purrr::reduce(complete.multiview.forecasts, dplyr::full_join)

saveRDS(complete.multiview.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.multiview.forecasts.RDS"))
```


### Simplex EDM

```{r seg_univariatesimplexcomplete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_simplexEDM.txt'
file.create(write.file)

complete.simplex.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  percent <- 0.9
  lib <- c(1, ceiling(percent*nrow(dd.bottle)))
  pred <- c(ceiling(percent*nrow(dd.bottle))+1, nrow(dd.bottle))
  
  df <- lapply(targets, function(target){
    ddtarget <- dd.bottle[,c("day",target)]
    E <- UniSimplex_determineE(ddtarget, target, maxE = 40, lib, pred)
    forecast <- Simplex(dataFrame = ddtarget, lib = lib, 
                        pred = pred, target = target,
                        columns = target, E = E)
    rmse <- sqrt(mean((forecast$Observations-forecast$Predictions)^2, na.rm = T))
    ddrmse <- data.frame(Target=target, RMSE=rmse, E=E)
    ddrmse
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = 1)

complete.simplex.forecasts <- purrr::reduce(complete.simplex.forecasts, dplyr::full_join)

saveRDS(complete.simplex.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.simplex.forecasts.RDS"))
```


### ARIMA

```{r seg_arima_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_ARIMA.txt'
file.create(write.file)

complete.arima.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  covariates <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  b.train <- dd.bottle[lib,]
  b.forecast <- dd.bottle[pred,] 
  
  ### Forecasting
  
  df <- lapply(targets, function(target){
    
    idx <- which(target==covariates)
    predictors <- covariates[-idx]
    m <- auto.arima(b.train[target],max.p=10,max.q=10,max.d=3, max.order = 15) #d=0, xreg = as.matrix(b.train[,predictors])
    # checkresiduals(m)
    forecast <- fitted(Arima(unlist(dd.bottle[,target]),model=m))[pred] #xreg = as.matrix(dd.bottle[,predictors])
    
    rmse <- sqrt(mean((unlist(b.forecast[target])-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
  
}, mc.cores = detectCores()-2) 

complete.arima.forecasts <- purrr::reduce(complete.arima.forecasts, dplyr::full_join)

saveRDS(complete.arima.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.arima.forecasts.RDS"))
```

### Recurrent neural networks

```{r seg_RNN_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
virtualenv_create("myenv")
use_virtualenv("myenv")
# set some parameters for our model
batch_size <- 64 # number of sequences to look at at one time during training
total_epochs <- 30 # how many times we'll look @ the whole dataset while training our model

set.seed(1231)

### Log file
write.file = 'seg_ForecastLog_completeTS_RNN.txt'
file.create(write.file)

complete.rnn.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  X_train <- as.matrix(dd.bottle[lib[-length(lib)], possible_predictors])
  X_test <- as.matrix(dd.bottle[pred[-length(pred)], possible_predictors])
  
  df <- lapply(targets, function(target){
    
    y_train <- as.matrix(dd.bottle[lib[-1], target])
    y_test <- as.matrix(dd.bottle[pred[-1], target])
    
    # initialize our model
    # install_keras()
    model <- keras_model_sequential() %>%
      layer_dense(units = 7, activation = 'relu', input_shape = length(possible_predictors), 
                  kernel_initializer=initializer_random_uniform(seed = 43)) %>% 
      layer_dense(units = 49, activation = 'relu',
                  kernel_initializer=initializer_random_uniform(seed = 23)) %>%
      layer_dense(units = 1, activation = 'linear',
                  kernel_initializer=initializer_random_uniform(seed = 55))
    
    model %>% compile(
      loss = 'mean_squared_error',
      optimizer = 'adam',
      metrics = c('mae')
    )
    
    # Actually train our model! This step will take a while
    trained_model <- model %>% fit(
      x = X_train, # sequence we're using for prediction 
      y = y_train, # sequence we're predicting
      batch_size = batch_size, # how many samples to pass to our model at a time
      epochs = total_epochs, # how many times we'll look @ the whole dataset
      verbose=0,
      validation_split = 0.1) # how much data to hold out for testing as we go along
    
    forecast <- data.frame(y = predict(model, x=X_test))
    
    rmse <- sqrt(mean(unlist(y_test-forecast)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rnn.forecasts <- purrr::reduce(complete.rnn.forecasts, dplyr::full_join)
saveRDS(complete.rnn.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rnn.forecasts.RDS"))
```


### Random forest

```{r seg_RF_forecast_complete, message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
set.seed(1231)

max_lag = 4
steps_ahead = 1

### Log file
write.file = 'seg_ForecastLog_completeTS_RF.txt'
file.create(write.file)

complete.rf.forecasts <- mclapply(complete.ts.dd.seg.bottle, function(dd.bottle){
  
  bot <- unique(dd.bottle$bottle)
  
  bottle_specs <- dd.bottle %>%
    summarize(bottle = unique(bottle),
              richness = unique(richness),
              composition = unique(composition),
              incubator = unique(incubator),
              realized_richness = unique(realized_richness),
              mean_richness = unique(mean_richness), 
              median_richness = unique(median_richness),
              light_treatment = unique(light_treatment))
  
  ### Log file
  write(paste0("Bottle: ", bot, " started at: ", Sys.time(), "\n"), write.file, append = T)
  bot.start.time <- Sys.time()
  
  ### Bottle specific preparations
  dd.bottle <- dd.bottle[,colSums(is.na(dd.bottle))<nrow(dd.bottle)] %>%
    ungroup()
  
  targets <- dd_list$densities.int.detrend.complete
  targets <- unlist(unique(targets[targets$bottle==bot & targets$ForecastTarget==T, "species"]))
  targets <- c(targets, additional_targets)
  targets <- gsub(" ", "_", targets)
  targets <- unname(targets)
  
  idx <- which(colnames(dd.bottle) %in% all_possible_predictors)
  possible_predictors <- unique(c(targets, colnames(dd.bottle)[idx]))
  possible_predictors <- possible_predictors[!(possible_predictors %in% c("Didinium_nasutum","Debris",
                                                                          "Digested_algae","Small_unidentified",
                                                                          "Algae"))]
  possible_predictors <- gsub(" ", "_", possible_predictors)
  
  percent <- 0.9
  lib <- 1:ceiling(percent*nrow(dd.bottle))
  pred <- (ceiling(percent*nrow(dd.bottle))+1):nrow(dd.bottle)
  pred <- pred - max_lag
  lib <- lib - max_lag
  lib <- lib[lib>0]
  pred <- pred[pred>0]

  df <- lapply(targets, function(target){
    
    df.emb <- dd.bottle[,target]
    for(l in 1:max_lag){
      predictor.lag <- lag(dd.bottle, steps_ahead + l - 1)
      colnames(predictor.lag) <- paste0(colnames(predictor.lag), "_", l)
      df.emb <- cbind(df.emb, predictor.lag)
    }
    df.emb <- df.emb %>% na.omit()
    
    # split into training and test data, then forecast
    
    test <- df.emb[pred,]
    train <- df.emb[lib,]
    
    y_train <- train[, 1] # the target
    X_train <- train[, -1] # everything but the target
    
    y_test <- test[, 1] # the target
    X_test <- test[, -1] # everything but the target
    
    # fit the model
    fit_rf <- randomForest(X_train, y_train)
    
    # predict using the test set
    p <- predict(fit_rf, X_test)
    forecasts_rf <- data.frame(observed=y_test,
                               predicted=p)
    
    rmse <- sqrt(mean((forecasts_rf$observed-forecasts_rf$predicted)^2, na.rm = T))
    
    data.frame(Target=target, RMSE=rmse)
  })
  
  df <- do.call("rbind", df)
  df$bottle <- bot
  df <- full_join(bottle_specs,df)
  
  ### Log file
  elapsed <- Sys.time() - bot.start.time
  elapsed <- paste(round(elapsed,2), attr(elapsed, "units"))
  write(paste0("Bottle: ", bot, " ended at: ", Sys.time(), "; Elapsed time: ", elapsed, "\n"), write.file, append = T)
  
  ## return 
  df
}, mc.cores=detectCores()-2)

complete.rf.forecasts <- purrr::reduce(complete.rf.forecasts, dplyr::full_join)
saveRDS(complete.rf.forecasts, file = here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rf.forecasts.RDS"))
```


## Post-forecasting data wrangling 

```{r forecast_post_processing, message=FALSE, warning=FALSE}
# based on segmented regression detrending

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/2_SegmentedAnalysis/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"


seg.complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                             complete.multiview.forecasts,
                                             complete.rnn.forecasts,
                                             complete.simplex.forecasts,
                                             complete.rf.forecasts), 
                                        dplyr::full_join)

# based on simple temporal regression detrending

complete.arima.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.arima.forecasts.RDS"))
complete.multiview.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.multiview.forecasts.RDS"))
complete.rnn.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rnn.forecasts.RDS"))
complete.simplex.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.simplex.forecasts.RDS"))
complete.rf.forecasts <- readRDS(here("Data/3_ForecastsResults/1_DetrendedLinearly/complete.rf.forecasts.RDS"))

complete.arima.forecasts$forecast.method <- "ARIMA"
complete.multiview.forecasts$forecast.method <- "Multiview EDM"
complete.rnn.forecasts$forecast.method <- "RNN"
complete.simplex.forecasts$forecast.method <- "Simplex EDM"
complete.rf.forecasts$forecast.method <- "Random Forest"

complete.forecasts <- purrr::reduce(list(complete.arima.forecasts,
                                         complete.multiview.forecasts,
                                         complete.rnn.forecasts,
                                         complete.rf.forecasts,
                                         complete.simplex.forecasts), 
                                    dplyr::full_join)
```


# Calculation of temporal stability and permutation entropy

```{r complete.CV.calc, message=FALSE, warning=FALSE, cache=FALSE}
biomass.complete.pe.cv <- dd_list$biomass.int.detrend.complete %>%
  group_by(bottle, composition, richness, incubator, light_treatment, realized_richness, mean_richness, median_richness) %>%
  summarize(biomass.pe.5 = PE(x = biomass, weighted = T, word_length = 5, tau = 1,
                              tie_method = "noise", noise_amount = 10),
            biomass.cv.res = sd(biomass.detrended.unstandardized, na.rm=T)/mean(biomass.original, na.rm=T),
            biomass.acf_lag5 = stats::acf(biomass, lag.max=5, plot=F, na.action = na.pass)$acf[6])

density.complete.pe.cv <- dd_list$densities.int.detrend.complete %>%
  group_by(species, bottle, composition, ForecastTarget, richness, incubator, method, light_treatment, realized_richness, mean_richness, median_richness) %>%
  summarize(density.pe.5 = PE(x = density, weighted = T, word_length = 5, tau = 1,
                              tie_method = "noise", noise_amount = 10),
            biomass.pe.5 = ifelse(all(is.na(biomass)), NA, PE(x = biomass, weighted = T, word_length = 5, tau = 1,
                                                              tie_method = "noise", noise_amount = 10)),
            biomass.cv.res = sd(biomass.detrended.unstandardized, na.rm=T)/mean(biomass.original, na.rm=T),
            density.cv.res = sd(density.detrended.unstandardized, na.rm=T)/mean(density.original, na.rm=T),
            biomass.acf_lag5 = stats::acf(biomass, lag.max=5, plot=F, na.action = na.pass)$acf[6],
            density.acf_lag5 = stats::acf(density, lag.max=5, plot=F, na.action = na.pass)$acf[6])

o2.complete.pe.cv <- dd_list$o2.int.detrend.complete %>%
  group_by(bottle, composition, light_treatment, richness, realized_richness, mean_richness, median_richness) %>%
  summarize(percent_o2.pe.5 = PE(x = percent_o2, weighted = T, word_length = 5, tau = 1,
                                 tie_method = "noise", noise_amount = 10),
            percent_o2.cv.res = sd(percent_o2.detrended.unstandardized, na.rm=T)/mean(percent_o2.original, na.rm=T),
            percent_o2.acf_lag5 = stats::acf(percent_o2, lag.max=5, plot=F, na.action = na.pass)$acf[6])

water_chem.complete.pe.cv <- dd_list$water_chem.int.detrend.complete %>%
  group_by(type, bottle, composition, light_treatment, richness, realized_richness, mean_richness, median_richness) %>%
  summarize(concentration.pe.5 = PE(x = concentration, weighted = T, word_length = 5, tau = 1,
                                    tie_method = "noise", noise_amount = 10),
            concentration.cv.res = sd(concentration.detrended.unstandardized, na.rm=T)/mean(concentration.original, na.rm=T),
            concentration.acf_lag5 = stats::acf(concentration, lag.max=5, plot=F, na.action = na.pass)$acf[6])
```


## Addition to forecast results

```{r mergeForecastsandTSMetrics, warning=FALSE, message=FALSE}
complete.forecasts <- complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))

seg.complete.forecasts <- seg.complete.forecasts %>%
  group_by(forecast.method) %>%
  mutate(realized_richness_center = realized_richness - round(mean(realized_richness)),
         median_richness_center = median_richness - (median(median_richness)),
         mean_richness_center = mean_richness - (mean(mean_richness)))


complete.forecasts.pe.cv <- full_join(complete.forecasts %>%
                                        dplyr::filter(Target == "Total_biomass"),
                                      biomass.complete.pe.cv %>%
                                        mutate(Target = "Total_biomass") %>%
                                        rename(cv.res=biomass.cv.res,
                                               acf_lag5 = biomass.acf_lag5,
                                               pe5=biomass.pe.5))

complete.forecasts.pe.cv2 <- full_join(complete.forecasts %>%
                                         dplyr::filter(Target != "Total_biomass",Target != "Mean_oxygen"),
                                       density.complete.pe.cv %>%
                                         ungroup() %>%
                                         mutate(Target=gsub(" ", "_", species)) %>%
                                         dplyr::select(-biomass.pe.5, -biomass.cv.res, -biomass.acf_lag5,
                                                       -species, -method, -ForecastTarget) %>%
                                         rename(cv.res=density.cv.res,
                                                acf_lag5 = density.acf_lag5,
                                                pe5=density.pe.5)) %>%
  drop_na(RMSE)

complete.forecasts.pe.cv3 <- full_join(complete.forecasts %>%
                                         dplyr::filter(Target == "Mean_oxygen"),
                                       o2.complete.pe.cv %>%
                                         mutate(Target = "Mean_oxygen") %>%
                                         rename(cv.res=percent_o2.cv.res,
                                                acf_lag5 = percent_o2.acf_lag5,
                                                pe5=percent_o2.pe.5))

complete.forecasts.pe.cv <- rbind(complete.forecasts.pe.cv,
                                  complete.forecasts.pe.cv2,
                                  complete.forecasts.pe.cv3)

complete.forecasts <- complete.forecasts.pe.cv %>%
  dplyr::select("bottle","richness","composition","incubator","realized_richness",
                "light_treatment","Target", "RMSE","forecast.method",
                "mean_richness","median_richness","realized_richness_center",
                "median_richness_center","mean_richness_center","pe5",
                "cv.res","acf_lag5") %>%
  mutate(cv.res.log10 = log10(cv.res))

seg.complete.forecasts <- seg.complete.forecasts %>%
  dplyr::select("bottle","richness","composition","incubator","realized_richness",
                "light_treatment","Target", "RMSE","forecast.method",
                "mean_richness","median_richness","realized_richness_center",
                "median_richness_center","mean_richness_center")
```


# Saving merged and processed results

```{r}
save(complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/complete.forecasts.RData"))
save(seg.complete.forecasts, 
     file = here("Data/3_ForecastsResults/3_MergedResults/seg.complete.forecasts.RData"))
``` 

